Carbonate chemistry at Ny-Ålesund

### 2021-02-01 awipev-CO2.Rmd
Rmd works again

---
title: "Carbonate chemistry at Ny-Ålesund: AWIPEV-CO~2~ Project" 
author: "Jean-Pierre Gattuso and Samir Alliouane"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  rmarkdown::
  rmarkdown::html_document:
    theme: paper
    number_sections: false

---
![ ](Images/AWIPEV-Logo.jpg){width=250px}

@@

#################### Function RMSE for model 2 regression root mean square error
lm2 <- function(data = NULL, x, y) {
  xs <- deparse(substitute(x))
  ys <- deparse(substitute(y))
#fit <- lmodel2(formula = y  ~ x, data = data, nperm = 99)
  print(y)
  #return(fit)

  lmwrap <- function(d,y,x) {
    ys <- deparse(substitute(y))
    xs <- deparse(substitute(x))
    f <- reformulate(xs,response=ys)
    return(lm(f,data=d))
}
mydata <- data.frame(X=1:10,Y=rnorm(10))
lmwrap(mydata,Y, X)

lm2(x = d_all$pco2_calc_fb, y = d_all$pco2_fb)
lm2(data = d_all, x=pco2_calc_fb, y=pco2_fb)

lmodel2(d_all$pco2_calc_fb ~ d_all$pco2_fb, nperm = 99)

  intercept_MA <- fit$regression.results[2,2]
  slope_MA <- fit$regression.results[2,3]
  intercept_OLS <- fit$regression.results[1,2]
  slope_OLS <- fit$regression.results[1,3]
  predicted <- intercept_OLS + df[[x]] * slope_OLS
  error <- x - predicted
  my_list <- list(fit = fit, rmse = sqrt(mean(error^2, na.rm=TRUE)))
  return(my_list)
  }

@@

## Impact of choice of constants
### Using mean values
```{r impact of constants, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis
# use mean values calculated for impacts of nutrients above
means <- d_all %>% 
  group_by(montht) %>% 
  summarize(mean_ct = mean(ct, na.rm = TRUE),
            mean_at_calc = mean(at_calc, na.rm = TRUE),
            mean_pco2_calc = mean(pco2_calc_fb, na.rm = TRUE),
            mean_sal = mean(sal_fb, na.rm = TRUE),
            mean_temp = mean(temp_fb, na.rm = TRUE)) %>% 
  ungroup()
mean_ct <-  mean(means$mean_ct)
mean_at_calc <-  mean(means$mean_at_calc)
mean_pco2_calc <-  mean(means$mean_pco2_calc)
mean_sal <-  mean(means$mean_sal)
mean_temp <-  mean(means$mean_temp)
# impacts using pco2 and AT
carb_l_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1

# impacts using AT-DIC pair
carb_l_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
const <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag:T, pH, pCO2, DIC, OmegaAragonite)
#kable(const, digits = c(0, 1, 1, 3, 1, 1, 2))
pander(const)
```
### Using all values pCO2 and at_calc
```{r impact of constants, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis
d_ls <- d_all %>% 
  drop_na(pco2_fb, at_calc, sal_fb, temp_fb)
carb_l <-
  carb(
    24,
    d_ls$pco2_fb,
    d_ls$at_calc * 1e-6,
    S = d_ls$sal_fb,
    T = d_ls$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "l",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
carb_s <-
  carb(
    24,
    d_ls$pco2_fb,
    d_ls$at_calc * 1e-6,
    S = d_ls$sal_fb,
    T = d_ls$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "s20",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
diff_ls <- as_tibble(carb_l - carb_s)
d_ls <- bind_cols(d_ls, diff_ls)

# plot
lm <- lm2(x = tmp$sal_insitu_ctd_9m, y = d_all$sal_fb)
p <-  ggplot(d_all,aes(x=sal_insitu_ctd_9m, y= sal_fb))+
  geom_point( aes(color=factor(monthd))) + 
  geom_abline(slope=1, intercept=0, linetype = "dashed")+
  geom_abline(aes(intercept = lm$fit$regression.results[2,2], slope = lm$fit$regression.results[2,3]),colour = "blue") +
  ylim(c(30.5,35.5)) +xlim(c(28,35.5)) +
  labs(title = paste("","\n Adj R2 = ", signif(lm$fit$rsquare, 3),
                     "; Intercept =", signif(lm$fit$regression.results[2,2], 3),
                     "; Slope =", signif(lm$fit$regression.results[2,3], 3),
                     "; P =", signif(lm$fit$P.param, 3),
                     "; RMSE = ", signif(lm$rmse, 3))) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 7)) + #
  coord_fixed(ratio = 1 )+
  labs( x="Salinity CTD 9 m", y="Salinity FB ", color = "Months") + theme_bw()

@@

The difference induces by the choice of the carbonic acid constants is large for pH, pCO2 and DIC and negligible for OmegaAragonite.

>>>>

---
title: "Carbonate chemistry at Ny-Ålesund" 
author: "Jean-Pierre Gattuso and Samir Alliouane"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  rmarkdown::html_document:
    theme: paper
    number_sections: false
---
![ ](Images/AWIPEV-Logo.jpg){width=250px}

@@

#################### Function RMSE for model 2 regression root mean square error
#test <- readr::read_csv(file="test_file")
lm2 <- function(data = NULL, x, y) {
  col_x <- eval(substitute(x), data)
  col_y <- eval(substitute(y), data)
  col_x <- x
  col_y <- y
fit <- lmodel2(col_y ~ col_x, nperm = 99)
  intercept_MA <- fit$regression.results[2,2]
  slope_MA <- fit$regression.results[2,3]
  intercept_OLS <- fit$regression.results[1,2]
  slope_OLS <- fit$regression.results[1,3]
  predicted <- intercept_OLS + col_x * slope_OLS
  error <- col_x - predicted
  my_list <- list(fit = print(fit), rmse = sqrt(mean(error^2, na.rm=TRUE)))
  return(my_list)
}
#z <- lm2(x = d_all$pco2_calc_fb, y = d_all$pco2_fb)
#z <- lmodel2(d_all$pco2_fb ~ d_all$pco2_calc_fb)
#z <- lm2(data = d_all, x = pco2_calc_fb, y = pco2_fb)

@@

## Impact of choice of constants
### Using mean values
```{r impact of constants 1, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis
# use mean values calculated for impacts of nutrients above
means <- d_all %>% 
  group_by(montht) %>% 
  summarize(mean_ct = mean(ct, na.rm = TRUE),
            mean_at_calc = mean(at_calc, na.rm = TRUE),
            mean_pco2_calc = mean(pco2_calc_fb, na.rm = TRUE),
            mean_sal = mean(sal_fb, na.rm = TRUE),
            mean_temp = mean(temp_fb, na.rm = TRUE)) %>% 
  ungroup()
mean_ct <-  mean(means$mean_ct)
mean_at_calc <-  mean(means$mean_at_calc)
mean_pco2_calc <-  mean(means$mean_pco2_calc)
mean_sal <-  mean(means$mean_sal)
mean_temp <-  mean(means$mean_temp)
# impacts using pco2 and AT
carb_l_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1

# impacts using AT-DIC pair
carb_l_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
const <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag:T, pH, pCO2, DIC, OmegaAragonite)
#kable(const, digits = c(0, 1, 1, 3, 1, 1, 2))
pander(const)
```
### Using all values pCO2 and at_calc

```{r impact of constants 2, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis
d_ls <- d_all %>% 
  drop_na(pco2_fb, at_calc, sal_fb, temp_fb)
carb_l <-
  carb(
    24,
    d_ls$pco2_fb,
    d_ls$at_calc * 1e-6,
    S = d_ls$sal_fb,
    T = d_ls$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "l",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
carb_s <-
  carb(
    24,
    d_ls$pco2_fb,
    d_ls$at_calc * 1e-6,
    S = d_ls$sal_fb,
    T = d_ls$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "s20",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
diff_ls <- as_tibble(carb_l - carb_s)
d_ls <- bind_cols(d_ls, diff_ls)

# plot
lm <- lm2(x = d_all$sal_insitu_ctd_9m, y = d_all$sal_fb)
p <-  ggplot(d_all,aes(x=sal_insitu_ctd_9m, y= sal_fb))+
  geom_point( aes(color=factor(monthd))) + 
  geom_abline(slope=1, intercept=0, linetype = "dashed")+
  geom_abline(aes(intercept = lm$fit$regression.results[2,2], slope = lm$fit$regression.results[2,3]),colour = "blue") +
  ylim(c(30.5,35.5)) +xlim(c(28,35.5)) +
  labs(title = paste("","\n Adj R2 = ", signif(lm$fit$rsquare, 3),
                     "; Intercept =", signif(lm$fit$regression.results[2,2], 3),
                     "; Slope =", signif(lm$fit$regression.results[2,3], 3),
                     "; P =", signif(lm$fit$P.param, 3),
                     "; RMSE = ", signif(lm$rmse, 3))) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 7)) + #
  coord_fixed(ratio = 1 )+
  labs( x="Salinity CTD 9 m", y="Salinity FB ", color = "Months") + theme_bw()

@@

The difference induced by the choice of the carbonic acid constants is large for pH, pCO2 and DIC and negligible for OmegaAragonite.
